Thesis: draft

### 2021-02-12 draft.tex
continued writing on real world social networks + introduction on opinion dynamics

\chapter{Opinion dynamics}
\label{chap3}
% talk about opinion dynamics, different models, resistance, etc

>>>>

\section{Real-world social networks}
The models discussed in the sections above, describe real-world networks in an idealized, simplified way. They may posses some of the characteristics of real-world networks, but they certainly do not always describe real-world networks appropriately. The question then arises: what are the main properties/characteristics of real-world networks, and more specifically of real-world social networks? Another important question that can be asked is whether or not there is a significant difference between off-line social networks and on-line social networks?\\
As already briefly discussed in the introduction, it is found that real-world networks often posses three characteristic properties. These are the two properties of the "small world phenomenon", namely a short average distance and a high clustering coefficient, and a heterogeneous or power law degree distribution \cite{RealWorld}. These properties are valid for both off-line and on-line social networks \cite{Zhang2014}. Lets discuss them in a bit more detail.\\
\textbf{Short average distance}\\
In the late 1960's Stanley Milgram performed a, what is now famous, experiment. He asked randomly selected people in Boston and Omaha to forward a letter to a distant target person. The letter could, however, only be sent to acquaintances, thought to be closer to the target person \cite{RealWorld}. The remarkable result of this experiment was that the average number of steps between the sender and the target person was only around six. This phenomenon is now often referred to as "six degrees of separation" \cite{RealWorld}.\\
This phenomenon is also observed in many other real-world networks: in most real-world networks it is possible to go from one node to another node through a number of edges that is small compared to the system size \cite{RealWorld}. Such networks usually have an average distance that depends logarithmically on the system size instead of linearly. A short average distance facilitates a fast transmission of information \cite{Zhang2014}.\\
\textbf{High clustering coefficient}\\
In real-world social networks, one often observes the pattern that two acquaintances/friends of an individual are also acquaintances/friends of each other \cite{RealWorld}. In graph terminology, this is translated into a high clustering coefficient (which quantifies a large number of triangles in the network) \cite{RealWorld}. This property is not only found in social networks, but also in a wide variety of different real-world networks \cite{RealWorld}.\\
\textbf{Heterogeneous or power law degree distribution}\\
Many real-world networks are found to have a power law degree distribution \cite{RealWorld}
	P(k) \sim k^{-\gamma}
where the exponent $\gamma$ often lies between 2 and 3. Networks with a power law degree distribution are also called heterogeneous or scale free networks. The term scale free refers to the fact that there is no characteristic scale in the network. Compared to the homogeneous network, where each node has a degree close to the average degree $\left<k\right>$, the heterogeneous network has many nodes with a low degree and a few nodes with a very high degree (these nodes are referred to as hubs). If the exponent $\gamma$ is smaller or equal to 2 ($\gamma \leqslant 2$), the average degree diverges; if $2 < \gamma < 3$, there is a finite average degree, but a diverging variance \cite{Newman2005}.\\
As said earlier, these three main characteristics are the same for on- and off-line social networks. Another property that is found in most on- and off-line networks is the appearance of a giant component, this is, most nodes belong to the same connected component \cite{Latapy}. They also often posses a modular/community structure \cite{Ferrara2012}\cite{McGlohon2011}. There are, however, some properties that are different between on-line and off-line networks as well. One of them might be the degree correlations/mixing pattern. It is commonly assumed and accepted that most real-life social networks have an assortative mixing pattern \cite{Hu2009}\cite{Zhang2014}. For a long time, it was thought that this is also valid for on-line social networks. However, recent research has shown that a lot of on-line social networks display disassortative mixture \cite{Hu2009}\cite{Zhang2014}. This discrepancy between real-life social networks and on-line social networks may be explained as follows. In real-life, `ordinary people' often want to be friends with celebrities, however, the celebrities usually prefer to stay within their own circles \cite{Hu2009}\cite{Zhang2014}. The `ordinary people' don't have access to these circles and we are left with the situation where `ordinary people' interact with `ordinary people' and celebrities hang out among other celebrities; hence, an assortative structure. In on-line social networks, `ordinary people' can easily connect with the celebrity and vice-versa, the celebrity wants to show his/her importance/influence by the number of fans \cite{Hu2009}\cite{Zhang2014}. This leads to a disassortative mixing pattern.
% also talk about size of groups that people can maintain?

\chapter{Opinion dynamics}
\label{chap3}
The study of opinion dynamics in a group of people, as already mentioned in the introduction, has two major layers: a network layer and an opinion dynamics layer. The network layer that describes the underlying structure and correlations between the agents/people is intensively described in the previous section. In this section we will focus on the second important layer, namely that of the opinion dynamics models. These models need to describe, in a simplified way, how opinions are formed and how they evolve in a group of people that interact in a certain way. Hence, these opinion dynamics are the dynamics that run on the underlying network structure. Several models have been proposed; here we will review some of the most important ones, with the main focus on the models used in this thesis. Some major attention will also be given to the content curation/filtering algorithms used by social media companies. This is because this thesis deals with opinion dynamics on social media, so the aspect of filtering cannot be ignored in the formation and evolution of opinions.
% talk about opinion dynamics, different models, resistance, etc
